---
title: "Single-cell Analysis of Host-Microbiome Interactions"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

<!-- badges: start -->
<!-- badges: end -->

## Installation
Here, we use `pak` to install package, you can also use `remotes`
```{r, eval=FALSE}
if (!requireNamespace("pak")) {
    install.packages("pak",
        repos = sprintf(
            "https://r-lib.github.io/p/pak/devel/%s/%s/%s",
            .Platform$pkgType, R.Version()$os, R.Version()$arch
        )
    )
}
```

`rsami` depends on [polars](https://rpolars.github.io/index.html), you can
install it by:
```{r, eval=FALSE}
pak::repo_add("https://rpolars.r-universe.dev")
pak::pkg_install("polars")
```

You can also install the development version of `polars` from
[Github](https://github.com/pola-rs/r-polars) 

In this way, you must install `rustc` to compile `polars`

You can install `rustc` with `rustup`:
```
curl --proto '=https' --tlsv1.2 -sSf https://sh.rustup.rs | sh
```
Then install the `polars` package:
```{r, eval=FALSE}
pak::pkg_install("pola-rs/r-polars")
```

You can install the development version of `rsahmi` from
[GitHub](https://github.com/Yunuuuu/rsahmi):

```{r, eval=FALSE}
pak::pkg_install("Yunuuuu/rsahmi")
```

You must also install [seqkit](https://bioinf.shenwei.me/seqkit/) and
[kraken2](https://github.com/DerrickWood/kraken2/wiki/Manual). 

## Cite
If you use this package, please cite:

 - Ghaddar B, Biswas A, Harris C, Omary MB, Carpizo DR, Blaser MJ, De S. Tumor
microbiome links cellular programs and immunity in pancreatic cancer. Cancer
Cell. 2022 Oct 10;40(10):1240-1253.e5. doi: 10.1016/j.ccell.2022.09.009. PMID:
36220074; PMCID: PMC9556978. 
 - Ghaddar B, Blaser MJ, De S. Denoising sparse microbial signals from
single-cell sequencing of mammalian host tissues. Nat Comput Sci. 2023
Sep;3(9):741-747. doi: 10.1038/s43588-023-00507-1. Epub 2023 Sep 18. PMID:
37946872; PMCID: PMC10634611.
 - Song Y, Peng Y, Qin C, Jiang S, Lin J, Lai S, Wu J, Ding M, Du Y, Yu L, Xu T.
Antibiotic use attenuates response to immune checkpoint blockade in urothelial
carcinoma via inhibiting CD74-MIF/COPA: revealing cross-talk between
anti-bacterial immunity and anti-tumor immunity through a tumor marker
prognostic study. Int J Surg. 2024 Jul 12. doi: 10.1097/JS9.0000000000001901.
Epub ahead of print. PMID: 38995167. 
