---
title: "Single-cell Analysis of Host-Microbiome Interactions"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/rsahmi)](https://CRAN.R-project.org/package=rsahmi)
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![](https://cranlogs.r-pkg.org/badges/rsahmi)](https://cran.r-project.org/package=rsahmi)
[![DOI](https://zenodo.org/badge/605180683.svg)](https://doi.org/10.5281/zenodo.15076675)
<!-- badges: end -->

A computational resource designed to accurately detect microbial nucleic acids
while filtering out contaminants and false-positive taxonomic assignments from
standard transcriptomic sequencing of mammalian tissues (See
[SAHMI](https://github.com/sjdlabgroup/SAHMI)). This implementation leverages
the `polars` package for fast and systematic microbial signal recovery and
denoising from host tissue genomic sequencing.

## Installation
You can install `rsahmi` from `CRAN` using:
```{r, eval=FALSE}
# install.packages("pak")
pak::pak("rsahmi")
```

Alternatively, install the development version from
[r-universe](https://yunuuuu.r-universe.dev/rsahmi) with:

```{r, eval=FALSE}
pak::repo_add("https://yunuuuu.r-universe.dev")
pak::pak("rsahmi")
```

or from [GitHub](https://github.com/Yunuuuu/rsahmi) with:
```{r, eval=FALSE}
pak::pak("Yunuuuu/rsahmi")
```

You must also install [seqkit](https://bioinf.shenwei.me/seqkit/) and
[kraken2](https://github.com/DerrickWood/kraken2/wiki/Manual).

Additionally, `rsahmi` relies on [polars](https://rpolars.github.io/index.html).
If you call any functions that require `polars`, you will be prompted to install
the package.

## Workflow
Prepare the output directory and input data

```{r, eval=FALSE}
odir <- "rsahmi" # Specify your output directory
# For 10x Genomic data, `fq1` only contain barcode and umi, but the official
# didn't give any information for this. In this way, I prefer using `umi-tools`
# to transform the `umi` into fq2 and then run `rsahmi` with only fq2.
fq1 <-      # Sequence file
fq2 <- NULL # can be `NULL`
if (dir.exists(odir)) dir.create(odir, recursive = TRUE)
```

### Taxonomic classifier 
Note: Replace `sample` with your actual sample name, and run the following code
for each of your samples. The function will generate the following files in the
`odir`, corresponding to the specified arguments::

- `kreport`: the `kraken2` report file.
- `koutput`: the `kraken2` output file.
- `classified_out`: the classified sequence file(s) from `kraken2`.
- `microbiome_koutput`: the microbiome output file extracted from `koutput`.
- `microbiome_reads`: the microbiome sequence reads file(s)

```{r, eval=FALSE}
kraken2_microbiome(
    fq1 = fq1,
    fq2 = fq2,
    # kraken2 database
    db = NULL,
    # Replace `sample` with your actual sample name. This will place the results
    # for each sample into its own directory.
    odir = file.path(odir, sample),
    # kraken2 command path
    kraken2 = NULL
)
```

**Following workflow will need all samples (multiple), you should run above command for your all samples, then proceed to the subsequent steps.** 

### Prepare datasets for signal denoising and taxa counting
```{r, eval=FALSE}
# paths: the output directory of all samples (from above steps)
paths <- fs::dir_ls(odir, type = "directory")
# this is the time-consuming step, so the internal will save the results in `odir`
sahmi_datasets <- lapply(paths, function(dir) {
    rsahmi::prep_dataset(
        fa1 = file.path(dir, "kraken_microbiome_reads.fa"),
        # if you have paired sequence, please also specify `fa2`,
        # !!! Also pay attention to the file name of `fa1` (add suffix `_1`) if
        #     you use paired reads.
        # fa2 = file.path(dir, "kraken_microbiome_reads_2.fa"),
        kraken_report = file.path(dir, "kraken_report.txt"),
        kraken_out = file.path(dir, "kraken_microbiome_output.txt"),
        odir = dir
    )
})
# If you have run above command, you can use following code
# to read the datasets
# sahmi_datasets <- lapply(paths, rsahmi::read_dataset)
```

### Identify contaminants
```{r, eval=FALSE}
library(polars) # rsahmi dependents on `polars`
library(ggplot2)
truly_microbe <- rsahmi::remove_contaminants(
    file.path(paths, "kraken_report.txt"),
    quantile = 0.99, exclusive = FALSE
)
microbe_for_plot <- attr(truly_microbe, "truly")[
    order(attr(truly_microbe, "pvalue")[attr(truly_microbe, "truly")])
]
microbe_for_plot <- microbe_for_plot[
    !microbe_for_plot %in% attr(truly_microbe, "exclusive")
]

truly_microbe_density <- ggplot(
    truly_microbe$filter(pl$col("taxid")$is_in(microbe_for_plot))$
        to_data_frame(),
    aes(rpmm),
) +
    geom_density(aes(fill = study), alpha = 0.5) +
    scale_x_log10() +
    facet_wrap(facets = vars(taxa), scales = "free") +
    theme(
        strip.clip = "off",
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "inside",
        legend.position.inside = c(1, 0),
        legend.justification.inside = c(1, 0)
    )
ggsave("truly_microbe.pdf",
    plot = truly_microbe_density,
    width = 8, height = 7L
)
```

### Sample level signal denoising
```{r, eval=FALSE}
slsd <- rsahmi::slsd(lapply(sahmi_datasets, `[[`, "kreport"))
real_taxids_slsd <- slsd$filter(
    pl$col("r1")$gt(0),
    pl$col("r2")$gt(0),
    pl$col("r3")$gt(0),
    pl$col("p1")$lt(0.05),
    pl$col("p2")$lt(0.05),
    pl$col("p3")$lt(0.05)
)$get_column("taxid")
```

### Filter contaminants
```{r, eval=FALSE}
umi_list <- lapply(sahmi_datasets, function(dataset) {
    # Barcode level signal denoising (barcode k-mer correlation test)
    blsd <- rsahmi::blsd(dataset$kmer)
    real_taxids <- blsd$filter(pl$col("padj")$lt(0.05))$get_column("taxid")
    # only keep taxids pass Sample level signal denoising
    real_taxids <- real_taxids$filter(real_taxids$is_in(real_taxids_slsd))
    # remove contaminants
    real_taxids <- real_taxids$filter(
        real_taxids$is_in(attr(truly_microbe, "truly", exact = TRUE))
    )
    # filter UMI data
    dataset$umi$filter(pl$col("taxid")$is_in(real_taxids))
})
```

### Taxa umi counting
```{r, eval=FALSE}
counts <- rsahmi::taxa_counts(umi_list, basename(names(umi_list)))
counts$write_csv(file.path(odir, "taxa_counts.txt"), separator = "\t")
```

## Cite
If you use this package, please cite:

- Song, Yuxuan PhD; Peng, Yun PhD; Qin, Caipeng PhD; Jiang, Shan PhD; Lin,
 Jiaxing PhD; Lai, Shicong MD; Wu, Jilin PhD; Ding, Mengting PhD; Du, Yiqing
 PhD*; Yu, Luping MD*; Xu, Tao MD*. Antibiotic use attenuates response to immune
 checkpoint blockade in urothelial carcinoma via inhibiting CD74-MIF/COPA:
 revealing cross-talk between anti-bacterial immunity and ant-itumor immunity.
 International Journal of Surgery 111(1):p 972-987, January 2025. | DOI:
 10.1097/JS9.0000000000001901.
- Ghaddar B, Blaser MJ, De S. Denoising sparse microbial signals from
single-cell sequencing of mammalian host tissues. Nat Comput Sci. 2023
Sep;3(9):741-747. doi: 10.1038/s43588-023-00507-1. Epub 2023 Sep 18. PMID:
37946872; PMCID: PMC10634611.

## sessionInfo
```{r}
sessionInfo()
```
