---
title: "Single-cell Analysis of Host-Microbiome Interactions"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

<!-- badges: start -->
<!-- badges: end -->

## Installation
Here, we use `pak` to install package, you can also use `remotes`
```{r, eval=FALSE}
if (!requireNamespace("pak")) {
    install.packages("pak",
        repos = sprintf(
            "https://r-lib.github.io/p/pak/devel/%s/%s/%s",
            .Platform$pkgType, R.Version()$os, R.Version()$arch
        )
    )
}
```

Then you can install `rsahmi`:

```{r, eval=FALSE}
pak::pkg_install("Yunuuuu/rsahmi")
```

You must also install [seqkit](https://bioinf.shenwei.me/seqkit/) and
[kraken2](https://github.com/DerrickWood/kraken2/wiki/Manual).

Additionally, `rsahmi` relies on [polars](https://rpolars.github.io/index.html).
If you call any functions that require `polars`, you will be prompted to install
the package.

## Cite
If you use this package, please cite:

 - Song, Yuxuan PhD; Peng, Yun PhD; Qin, Caipeng PhD; Jiang, Shan PhD; Lin,
 Jiaxing PhD; Lai, Shicong MD; Wu, Jilin PhD; Ding, Mengting PhD; Du, Yiqing
 PhD*; Yu, Luping MD*; Xu, Tao MD*. Antibiotic use attenuates response to immune
 checkpoint blockade in urothelial carcinoma via inhibiting CD74-MIF/COPA:
 revealing cross-talk between anti-bacterial immunity and ant-itumor immunity.
 International Journal of Surgery 111(1):p 972-987, January 2025. | DOI:
 10.1097/JS9.0000000000001901.
 - Ghaddar B, Biswas A, Harris C, Omary MB, Carpizo DR, Blaser MJ, De S. Tumor
microbiome links cellular programs and immunity in pancreatic cancer. Cancer
Cell. 2022 Oct 10;40(10):1240-1253.e5. doi: 10.1016/j.ccell.2022.09.009. PMID:
36220074; PMCID: PMC9556978.
 - Ghaddar B, Blaser MJ, De S. Denoising sparse microbial signals from
single-cell sequencing of mammalian host tissues. Nat Comput Sci. 2023
Sep;3(9):741-747. doi: 10.1038/s43588-023-00507-1. Epub 2023 Sep 18. PMID:
37946872; PMCID: PMC10634611.
